---
title: "EC349 Assignment R Markdown"
author: "Lina Tang"
date: "2023-12-02"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Libraries, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(tidyverse)
library(caret)
library(hexbin) 
library(jsonlite)
library(magrittr)
library(dplyr)
library(glmnet)
library(ggplot2)
library(tree)
library(rpart)
library(rpart.plot)
```

```{r message=FALSE, warning=FALSE}
getwd()
```

```{r message=FALSE, warning=FALSE}
setwd("C:/Users/123/Desktop/EC349 Assignment")
```

```{r load data, message=FALSE, warning=FALSE}
cat("\014")  
rm(list=ls())

load("C:/Users/123/Desktop/EC349 Assignment/yelp_review_small.Rda")
load("C:/Users/123/Desktop/EC349 Assignment/yelp_user_small.Rda")
```

```{r inspect data - review_data_small, message=FALSE, warning=FALSE}
head(review_data_small)
str(review_data_small)
summary(review_data_small)
dim(review_data_small)
colnames(review_data_small)
```

In the review sample, review_id, text and date will not be included as stars prediction factors because to their specificity and difficulty in definition.

```{r inspect data - user_data_small, message=FALSE, warning=FALSE}
head(user_data_small)
str(user_data_small)
summary(user_data_small)
dim(user_data_small)
colnames(user_data_small)
```

Meanwhile, in the user sample, name, yelping_since and elite will not be involved. From the brief inspection of data, the presence of extreme values implies the sample suffers from outlier problems. Initially, we will join the two samples based on the user_id and thereafter do data cleansing.

```{r merge data by "user_id", echo=TRUE, warning=FALSE, message=FALSE}
df <- review_data_small %>% left_join (user_data_small, by = "user_id")
str(df)
summary(df)
df$stars <- as.numeric(df$stars)
```

```{r  deal with missing values - delete, message=FALSE, warning=FALSE}
sapply(df, function(x) sum(is.na(x)))
df = df[complete.cases(df), ]
str(df)
summary(df)
```

While stars may appear to be a categorical variable, R Studio does not recognize factors as a response variable. All main variables will be designated as numerical values in the regression analysis but some of them will be placed as a factor in the decision tree. The combined sample has a significant amount of missing data. Following Zhang's (2015) approach, the missing values are excluded from the data. While scholars (Emmanuel et al., 2021) may argue the presence of bias, it is important to note that R Studio does not handle missing values in either regression or decision tree analyses.

```{r set all main variables as numeric, message=FALSE, warning=FALSE}
df$useful.x <- as.numeric(df$useful.x)
df$funny.x <- as.numeric(df$funny.x)
df$cool.x <- as.numeric(df$cool.x)
df$funny.y <- as.numeric(df$funny.y)
df$useful.y <- as.numeric(df$useful.y)
df$fans <- as.numeric(df$fans)
df$average_stars <- as.numeric(df$average_stars)
df$compliment_cool <- as.numeric(df$compliment_cool)
df$compliment_more <- as.numeric(df$compliment_more)
df$compliment_hot <- as.numeric(df$compliment_hot)
df$compliment_profile <- as.numeric(df$compliment_profile)
df$compliment_note <- as.numeric(df$compliment_note)
df$compliment_cute <- as.numeric(df$compliment_cute)
df$compliment_list <- as.numeric(df$compliment_list)
df$compliment_funny <- as.numeric(df$compliment_funny)
df$compliment_plain <- as.numeric(df$compliment_plain)
df$compliment_writer <- as.numeric(df$compliment_writer)
df$compliment_photos <- as.numeric(df$compliment_photos)
```

```{r deal with outliers, message=FALSE, warning=FALSE}
hist(df$useful.x, breaks=10)
hist(df$funny.x, breaks=10)
hist(df$cool.x, breaks=10)
df <- subset(df, useful.x < 6 & funny.x < 6 & cool.x < 6)
```

Without data variables description, review data involves the up-to-date evaluation to the restaurant. By contrast, user data involves users' previous accumulative reviews. In general, human or mechanical mistake, respondents' reluctance to answer specific questions, study dropout, merging unrelated data, and errors due to malfunctioning equipment are common causes of missing values (Emmanuel et al., 2021). The histograms demonstrated that a majority data of user.x, funny.x and cool.x are below 50. As stars are clearly categorical values ranging from 1 to 5, values for these variables are restricted to 5, to exclude outliers.

```{r split data into training and test samples, message=FALSE, warning=FALSE}
set.seed(1)
df$stars <- as.numeric(df$stars)
training.samples <- df$stars %>%
  createDataPartition(p = 0.8, list = FALSE)
train.data <- df[training.samples, ]
test.data <- df[-training.samples, ]
```

Data is randomly split into training and test samples, with probability of 0.8 and 0.2 respectively.

```{r Build model1, message=FALSE, warning=FALSE}
model1 <- lm(as.numeric(stars) ~ useful.x + funny.x + cool.x, data = train.data)
summary(model1)
plot(model1)
```

Upon observation, it is apparent that the factors of usefulness, humor, and attractiveness have a substantial impact on the rating of stars. The coefficient of R-squared (Wright, 1921) is the fraction of dependent variable variance predicted by independent factors. However, the capability of the model to account for the variations is inadequate, with an approximate R-squared of 7.25%. Among the other graphs, there exist numerous outliers. The relationship between the variables may be somewhat non-linear, as suggested by the non-linear red line. The presence of an S-shape in the Q-Q plot indicates that the data may deviate from a normal distribution.

```{r Prediction using model1, message=FALSE, warning=FALSE}
predictions_m1 <- predict(model1, newdata = test.data)
compare <- data.frame(actual = as.numeric(test.data$stars),
                      predicted = predictions_m1)
OLS_MSE_m1<- mean((predictions_m1 - as.numeric(test.data$stars)) ^ 2) #Note how it outperforms OLS
OLS_MSE_m1 #2.001981
sqrt(OLS_MSE_m1) #1.414914
```

```{r Model2: with more variables, message=FALSE, warning=FALSE}
model2 <- lm(as.numeric(stars) ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_hot + compliment_more + compliment_profile + compliment_cute + compliment_list + compliment_note + compliment_plain + compliment_cool + compliment_funny + compliment_writer + compliment_photos, data = train.data)
summary(model2)
plot(model2)
```

Clearly, compared to model 1, model 2, which includes additional variables, demonstrates a superior ability to account for the variation in error, as evidenced by a R^2 value of 37.67%. Model 2 appears to exhibit a normal distribution, as indicated by the updated Q-Q plot. In addition, these insignificant variables will be eliminated in order to create a more accurately fitting model.

```{r Prediction using model2, message=FALSE, warning=FALSE}
predictions_m2 <- model2 %>% predict(test.data)
compare <- data.frame(actual = as.numeric(test.data$stars),
                      predicted = predictions_m2)
OLS_MSE_m2<- mean((predictions_m2 - as.numeric(test.data$stars)) ^ 2) #Note how it outperforms OLS
OLS_MSE_m2 #1.332553
sqrt(OLS_MSE_m2) #1.154363
```

The RMSE of OLS linear regression model 2 is 1.154363.

```{r Model3: exclude insignificant affected variables, message=FALSE, warning=FALSE}
model3 <- lm(as.numeric(stars) ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_profile + compliment_more + compliment_list + compliment_writer + compliment_plain + compliment_note, data = train.data)
summary(model3)
plot(model3)
```


```{r Prediction using model3, message=FALSE, warning=FALSE}
predictions_m3 <- model3 %>% predict(test.data)
compare <- data.frame(actual = as.numeric(test.data$stars),
                      predicted = predictions_m3)
OLS_MSE_m3<- mean((predictions_m3 - as.numeric(test.data$stars)) ^ 2) #Note how it outperforms OLS
OLS_MSE_m3 #1.332569
sqrt(OLS_MSE_m3) #1.15437
```

The RMSE of OLS linear regression model 3 is 1.15437, which is exactly same as model 2. Therefore, we will focus on model 2 in the later analysis. 

```{r warning=FALSE, message=FALSE, include=FALSE}
library(sandwich)
library(lmtest)
coeftest(model3, vcov = vcovHC(model3, type="HC3"))
```

We will first take elastic nets to run ridge and LASSO regression respectively to make prediction on stars, and then decision tree.

```{r Ridge on Train Dataset, warning=FALSE, message=FALSE}
library(glmnet)
library(ggplot2)

set.seed(1)
train_y <- train.data$stars
train_x <- data.matrix(train.data[, c('useful.x', 'funny.x', 'cool.x', 'useful.y', 'funny.y','cool.y', 'average_stars', 'compliment_hot', 'review_count', 'compliment_more', 'compliment_profile', 'compliment_cute', 'compliment_list', 'compliment_note', 'compliment_plain', 'compliment_cool', 'compliment_funny', 'compliment_writer', 'compliment_photos')])

test_y <- test.data$stars
test_x <- data.matrix(test.data[, c('useful.x', 'funny.x', 'cool.x', 'useful.y', 'funny.y','cool.y', 'average_stars', 'compliment_hot', 'review_count', 'compliment_more', 'compliment_profile', 'compliment_cute', 'compliment_list', 'compliment_note', 'compliment_plain', 'compliment_cool', 'compliment_funny', 'compliment_writer', 'compliment_photos')])

ridge.mod <- glmnet(as.matrix(train_x), as.matrix(train_y), alpha = 0, lambda = 3, thresh = 1e-12)
summary(ridge.mod)
```

```{r message=FALSE, warning=FALSE}
#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(as.matrix(train_x), as.matrix(train_y), alpha = 0, nfolds = 3)

#find optimal lambda value that minimizes test MSE
lambda_ridge_cv <- cv_model$lambda.min
lambda_ridge_cv #0.08563297

#find optimal lambda value that minimizes test MSE
plot(cv_model)
```

Through cross-validation, the optimal lambda value is 0.08563297.

```{r message=FALSE, warning=FALSE}
#find coefficients of best model
best_model <- glmnet(as.matrix(train_x), as.matrix(train_y), alpha = 0, lambda = lambda_ridge_cv)
coef(best_model)
```

```{r message=FALSE, warning=FALSE}
#use fitted best model to make predictions
y_predicted <- predict(ridge.mod, s = lambda_ridge_cv, newx = test_x)

#find SST and SSE
sst <- sum((test_y - mean(test_y))^2)
sse <- sum((y_predicted - test_y)^2)

#find R-Squared
rsq <- 1 - sse/sst
rsq #0.2098919
```

```{r MSE for ridge model, message=FALSE, warning=FALSE}
ridge_MSE<- mean((y_predicted - test_y) ^ 2) #Note how it outperforms OLS
ridge_MSE #1.711681
sqrt(ridge_MSE) #1.308312
```

Surprisingly, the ridge model does not outperform OLS linear regression model 2, with higher RMSE.

```{r LASSO Model, message=FALSE, warning=FALSE}
cv.out2 <- cv.glmnet(as.matrix(train_x), as.matrix(train_y), alpha = 1, nfolds = 3)
plot(cv.out2)
lambda_LASSO_cv <- cv.out2$lambda.min #cross-validation is the lambda minimising empirical MSE in training data

#Re-Estimate Ridge with lambda chosen by Cross validation
LASSO.mod<-glmnet(train_x, train_y, alpha = 1, lambda = lambda_LASSO_cv, thresh = 1e-12)
coef(LASSO.mod) #note that some parameter estimates are set to 0 --> Model selection!

#Fit on Test Data
LASSO.pred <- predict(LASSO.mod, s = lambda_LASSO_cv, newx = as.matrix(test_x))
LASSO_MSE<- mean((LASSO.pred - test_y) ^ 2) #Note how it outperforms OLS
LASSO_MSE #1.332983
sqrt(LASSO_MSE) #1.154549
```

Relative standard error (RMSE) for LASSO regression and OLS model 2 are nearly same when using a similar set of variables. When it comes to star predictions, OLS regression model 2 outperforms LASSO regression in this case, but the ridge model was less effective. Although the stars are on a scale from 1 to 5, every single model has an MSE greater than 1, which is a relatively high value. Decision tree analysis allows for more flexibility in the specification of the relationship between variables; Thus, this technique will be used to forecast stars in the following project.

```{r Decision Tree, message=FALSE, warning=FALSE}
#What happens when you include all variables?
mod1 <- glm(stars ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_hot + compliment_more + compliment_profile + compliment_cute + compliment_list + compliment_note + compliment_plain + compliment_cool + compliment_funny + compliment_writer + compliment_photos, data = train.data)

#Review the results
coef(mod1) 
summary(mod1)

#Predicted Values for Test Data based on the model estimates
mod1_predict<-predict(mod1, newdata = test.data)
```

```{r message=FALSE, warning=FALSE}
#What happens when you exclude insignificant variables?
mod2 <- glm(stars ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_more + compliment_profile + compliment_list + compliment_note + compliment_plain + compliment_writer, data = train.data)

#Review the results
coef(mod2) 
summary(mod2)
```

Classification tree will be implemented since stars were declared to be categorical variables.

```{r Classification Tree, message=FALSE, warning=FALSE}
library(tree)
set.seed(1)
train.data$stars <- as.factor(train.data$stars)
test.data$stars <- as.factor(test.data$stars)
train.data$useful.x <- as.factor(train.data$useful.x)
test.data$useful.x <- as.factor(test.data$useful.x)
train.data$funny.x <- as.factor(train.data$funny.x)
test.data$funny.x <- as.factor(test.data$funny.x)
train.data$cool.x <- as.factor(train.data$cool.x)
test.data$cool.x <- as.factor(test.data$cool.x)

#With tree library
tree1<-tree(stars ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_more + compliment_profile + compliment_list + compliment_note + compliment_plain + compliment_writer, data = train.data)

#partition graph
partition.tree(tree1) 
points(train.data[, c("useful.x", "funny.x", "cool.x", "useful.y", "funny.y",   "cool.y", "average_stars", "fans", "review_count", "compliment_more",  "compliment_profile", "compliment_list", "compliment_note", "compliment_plain", "compliment_writer")], cex=.4)

plot(tree1)
text(tree1, pretty = 0)
title(main = "Unpruned Classification Tree")
```

```{r Classification Tree without restrictions, message=FALSE, warning=FALSE}
library(rpart)
rpart_tree<-rpart(stars ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_more + compliment_profile + compliment_list + compliment_note + compliment_plain + compliment_writer, data = train.data)
rpart.plot(rpart_tree)
```

Predicting stars at only 1 and 5, the decision tree is heavily influenced solely by average stars when minimum split and bucket limitations are not imposed.

```{r Classification Tree with restrictions, message=FALSE, warning=FALSE}
rpart_tree2<-rpart(stars ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_more + compliment_profile + compliment_list + compliment_note + compliment_plain + compliment_writer, data = train.data, control = rpart.control(maxdepth = 10, minsplit = 5, minbucket = 10000, cp = 0))
print(rpart_tree2)
rpart.plot(rpart_tree2, cex = 0.5)

#importance
rpart_tree2$variable.importance
```

In an effort to mitigate the problem of overfitting in the tree, the minimum bucket size is restricted to 10000 and the split is limited to 5. Simultaneously, the maximum depth is capped to 10. The revised classification tree incorporates both the number of reviews and the average star rating, resulting in more adaptable forecasts. 

While decision trees are conceptually uncomplicated and easy to execute and visualize, small perturbations in the data can have significant impacts on the resulting tree. Hence, there might be a considerable amount of variance and inadequate accuracy in making predictions. We will apply several techniques including bagging, boosting, and random forest to enhance the accuracy of our predictions. Bagging is an approximately unbiased model. In random forests, the concept of bootstrap aggregation is utilized in conjunction with random trees. Boosting lets trees grow in a way that gets rid of bias.

```{r Bagging, message=FALSE, warning=FALSE}
library(ipred)      
set.seed(1)      

#fit the bagged model
bag <- bagging(stars ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_more + compliment_profile + compliment_list + compliment_note + compliment_plain + compliment_writer, data = train.data, nbagg = 100,   
  coob = TRUE, control = rpart.control(minsplit = 3, cp = 0.1)
)

#display fitted bagged model
bag
```

The rate of misclassification error of bagging is relatively high, with an out-of-bag estimate of 0.4784. Specifically, with 100 bootstrap replications, the bagging classification trees accurately forecast star ratings 52.16% of the time. Inadequate modeling and improper execution of bagging can lead to significant bias, resulting in under-fitting. Then we will attempt random forests.

```{r Random Forest, message=FALSE, warning=FALSE}
set.seed(1)
library(randomForest) #randomforest
model_RF <- randomForest(stars ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_more + compliment_profile + compliment_list + compliment_note + compliment_plain + compliment_writer, data = train.data, ntree=200)
pred_RF_test = predict(model_RF, test.data)
mean(model_RF[["err.rate"]]) #0.5907672 for 200 trees.
```

Random forests do not perform well, with relatively high error rate of 0.5907672 for 200 trees.

```{r Boosting, message=FALSE, warning=FALSE}
library(adabag) #AdaBoost
# train a model using our training data
model_adaboost <- boosting(stars ~ useful.x + funny.x + cool.x + useful.y + funny.y + cool.y + average_stars + fans + review_count + compliment_more + compliment_profile + compliment_list + compliment_note + compliment_plain + compliment_writer, data = train.data, boos=TRUE, mfinal=50)
summary(model_adaboost)

# Load data for Regression (this one is stored on R)
#use model to make predictions on test data
pred_ada_test = predict(model_adaboost, test.data)

# Returns the prediction values of test data along with the confusion matrix
pred_ada_test
```

Unlike bagging averaging distinct training samples, boosting focuses on averaging different models. Boosting exhibits a reduced error rate of 0.4612786, surpassing the performance of bagging. Although bagging and boosting algorithms outperform classification trees, the resulting models cannot be interpretative in the form of trees.

